---
title: "Data Science"
author: "Maggie Sweitzer"
format: html
editor: visual
---

## Answers to Questions

### What do you think being a data scientist is about?

Data science combines skills in engineering, analysis, and modeling/inference to empirically derive answers to specific production-oriented questions from large volumes of data.

### What differences/similarities do you see between data scientists and statisticians?

According to Anuj Khandelwal, as described on the Indeed Career Guide (July 13, 2023), data scientists and statisticians both require the following skills:

-   Technical skills
    -   expertise in statistical analysis
    -   programming knowledge
    -   data visualization skills
    -   strong mathematical background
    -   knowledge of machine learning
    
-   Soft skills
    -   problem-solving
    -   communication
    -   attention to detail

In terms of differences, statisticians are likely to be more model or hypothesis-driven, whereas data scientists are more data-driven. The focus on data-driven analysis means that data scientists are likely to more heavily rely on coding languages such as Python to perform more complex data integration across large-scale data sets, whereas statisticians may work with a more circumscribed data set. In addition, data scientists are more likely to come from a broader range of educational backgrounds, including computer science or engineering, whereas statisticians often have specific training in statistics and/or math.

### How do you view yourself in relation to these two areas?

My area of study is very much at the intersection of these two areas. In my research I combine behavioral pharmacology and functional neuroimaging to answer questions about brain mechanisms underlying vulnerability to addiction. Because of the behavioral pharmacology component, I am often collecting data myself, sometimes longitudinally, and investing heavily in each participant. Typically the statistical models are defined in the grant, before funding is awarded and long before data collection begins. At the same time, there is a tremendous push in the field to utilize existing large-scale, multisite fMRI data sets to facilitate innovative, data-driven discoveries.

As I see it, there are a huge number of newly minted fMRI researchers entering the field with vastly more coding and machine learning experience than I will ever have. Therefore, in terms of data science, I would like to understand enough to ask the right questions and know when I might want to collaborate with the engineering/computer science experts. In the meantime, even though machine learning is currently all the rage, I think there will always be a need for traditional statistical modeling in psychiatry research, and that is a better fit for my background and interests. (But I do still want to learn Python).

## Running R Code

In this section, we first create an object y to store the results of the density function of Sepal.Length from the Iris data set:

```{r}
y <- density(iris$Sepal.Length)
```

Now, we will examine the class, type, and structure of the object y

```{r}
class(y)
typeof(y)
str(y)
```

Finally, we will use the plot function to visualize y

```{r}
#| echo: false

plot(y)
```
